### 2017-10-31 submission/manuscript.Rmd
update methods

***Study Analysis Overview:***

>>>>

***Study Analysis Overview:***
Alpha diversity was first assessed for differences between controls and adenoma versus cancer and controls versus adenoma. We analyzed the data using standard T-tests, ANOVA, linear mixed-effect models, and relative risk. Next, four specific CRC-associated genera (*Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, and *Porphyromonas*) were assessed for differences in relative risk. We then built Random Forest models based on all genera or the select CRC-associated genera. The models were trained on one study then tested on the remaining studies for every study. The data was split between feces and tissue samples. Within the tissue groups the data was further divided between matched and unmatched tissue samples. Both prediction for adenoma and carcinoma were tested. This same approach was then applied at the OTU level with the exception that instead of testing on the other studies a 10-fold cross validation was utilized and 100 different models were created based on random 80/20 splitting of the data to generate a range of expected AUCs. The power of each study was assessed for and effect size ranging from 1% to 30%. An estimated sample n for these effect sizes was also generated based on 80% power.

### 2017-10-31 submission/manuscript.tex
update methods

\textbf{\emph{Statistical Analysis.}} All statistical analysis after
sequence processing utilized the R software package (v3.4.2). For the
alpha diversity analysis values were power transformed using the
rcompanion (v1.10.1) package and then Z-score normalized using the car
(v2.1.5) package. Testing for alpha diversity differences utilized both
T-tests and ANOVA. Linear mixed-effect models were created using the
lme4 (v1.1.14) package to correct for both study and variable region
effect in the diversity measures when analyzing colorectal cancer
groups. Relative Risk was analyzed using both the epiR (v0.9.87) and
metafor (v2.0.0) packages. Relative risk significance testing utilized
the chi-squred test. Beta-diversity differences utilized a Bray-Curtis
distance matrix and PERMANOVA executed with the vegan (v2.4.4) package.
Random Forest models were built using both the caret (v6.0.77) and
randomForest (v4.6.12) packages. Random Forest testing of the obtained
AUC versus a random model AUC utilized T-tests. Power analysis and
estimations were made using the pwr (v1.2.1) and statmod (v1.4.30)
packages. All figures were created using both ggplot2 (v2.2.1) and
gridExtra (v2.3) packages.

\textbf{\emph{Reproducible Methods.}} The code and analysis can be found
here \url{https://github.com/SchlossLab/Sze_CRCMetaAnalysis_XXXX_2016}.
Unless mentioned otherwise the accession number for the raw sequences
for the studies used in this analysis can be found directly in the
respective batch file, on the GitHub repository or in the original
manuscript.

>>>>

\textbf{\emph{Statistical Analysis:}} All statistical analysis after
sequence processing utilized the R software package (v3.4.2). For the
alpha diversity analysis values were power transformed using the
rcompanion (v1.10.1) package and then Z-score normalized using the car
(v2.1.5) package. Testing for alpha diversity differences utilized both
T-tests and ANOVA. Linear mixed-effect models were created using the
lme4 (v1.1.14) package to correct for both study and variable region
effect in the diversity measures when analyzing colorectal cancer
groups. Relative Risk was analyzed using both the epiR (v0.9.87) and
metafor (v2.0.0) packages. Relative risk significance testing utilized
the chi-squred test. Beta-diversity differences utilized a Bray-Curtis
distance matrix and PERMANOVA executed with the vegan (v2.4.4) package.
Random Forest models were built using both the caret (v6.0.77) and
randomForest (v4.6.12) packages. Random Forest testing of the obtained
AUC versus a random model AUC utilized T-tests. Power analysis and
estimations were made using the pwr (v1.2.1) and statmod (v1.4.30)
packages. All figures were created using both ggplot2 (v2.2.1) and
gridExtra (v2.3) packages.

\textbf{\emph{Study Analysis Overview:}} Alpha diversity was first
assessed for differences between controls and adenoma versus cancer and
controls versus adenoma. We analyzed the data using standard T-tests,
ANOVA, linear mixed-effect models, and relative risk. Next, four
specific CRC-associated genera (\emph{Fusobacterium}, \emph{Parvimonas},
\emph{Peptostreptococcus}, and \emph{Porphyromonas}) were assessed for
differences in relative risk. We then built Random Forest models based
on all genera or the select CRC-associated genera. The models were
trained on one study then tested on the remaining studies for every
study. The data was split between feces and tissue samples. Within the
tissue groups the data was further divided between matched and unmatched
tissue samples. Both prediction for adenoma and carcinoma were tested.
This same approach was then applied at the OTU level with the exception
that instead of testing on the other studies a 10-fold cross validation
was utilized and 100 different models were created based on random 80/20
splitting of the data to generate a range of expected AUCs. The power of
each study was assessed for and effect size ranging from 1\% to 30\%. An
estimated sample n for these effect sizes was also generated based on
80\% power.

\textbf{\emph{Reproducible Methods:}} The code and analysis can be found
here \url{https://github.com/SchlossLab/Sze_CRCMetaAnalysis_XXXX_2016}.
Unless mentioned otherwise the accession number for the raw sequences
for the studies used in this analysis can be found directly in the
respective batch file, on the GitHub repository or in the original
manuscript.
